---
title: "Studying data"
author: "Melina Ribaud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Some real data sets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

In this vignette, we apply the functions of the package *olsgasp* to a real case example. This example is an extract from the methylation dataset of the article **mettre ref**. 

## Install and load package 

### Install packages
```{r, echo=TRUE, eval=FALSE}
install.packages("olsgasp")
install.packages("knitr")
install.packages("ggplot2")
install.packages("ggrepel")
install.packages("ggthemes")
install.packages("tidyr")
```

Note that *ggplot2*, *ggrepel* and *ggthemes* to plot graphics.

### Load packages
```{r, echo=TRUE, results='asis', warning=FALSE}
suppressPackageStartupMessages(library(olsgasp))
library(knitr)
library(ggplot2)
library(ggrepel)
library(ggthemes)
library(DiceEval)
library(tidyr)
```

## Load data

Load the dataset. 

```{r}
data(data_methyl)
Y = data_methyl$Y
sites = data_methyl$sites
X = data_methyl$X
colnames(X) = c("ES-cell","primary-cell")
K = nrow(Y)
N = ncol(Y)
```


## Simule missing data 

The missing values are located from sample 2 to 8 and 10 to 15 for the sites 20 to 100 and 120 to 150.  

```{r}
k_star = (1:K)[-c(1,9)]
n_star = c(20:100,120:150)
Y_obs = Y
Y_obs[k_star,n_star] = NA
```


## Initialize, fit and pred

The model is initialized with the function *svd_olsgasp*, fitted with *fit_olsgasp* and the missing values are predicted with *pred_olsgasp*. 


```{r}
obj_olsgasp = svd_olsgasp(Y_obs,sites,X,tol_eig = 1e-6)
obj_olsgasp = fit_olsgasp(obj_olsgasp)
Y_pred = pred_olsgasp(obj_olsgasp)
```

## Quality criteria

The Quality criteria are calculated only on the missing values. 

```{r}
ind_na = is.na(Y_obs)
R2(Y[ind_na],Y_pred[ind_na])
RMSE(Y[ind_na],Y_pred[ind_na])
```

## Visualize results

We choose two samples to visualize the true values of the processes and the predicted values. 

```{r}
proc_a = 2
proc_b = 10
```



```{r fig.width=5.5}
colorTitle = "black"
sizeTitle = 15
formeTitle = "bold.italic"
colorAxe = "black"
sizeAxe = 10
formeAxe = "bold"
textSize = 15
Title = "Two samples"
low = "#349be8"
high = "#cc0000"
point_size = 2 
size_point_graph = c(2,3)
x = X

df = data.table(sites = sites,
                  Y_A = Y[proc_a, ],
                  Y_A_obs = Y_pred[proc_a, ],
                  Y_B = Y[proc_b, ],
                  Y_B_obs = Y_pred[proc_b, ])
  
  
df = data.table(df %>% pivot_longer(!sites, names_to = "origine", values_to = "values"))
df = data.table(df, Type = factor(rep(c("Pred","True"),nrow(df)/2),labels = c("True","Pred")), 
                  X = as.factor(rep(rep(c(x[proc_a],x[proc_b]),each=2),nrow(df)/4)))
    
  p = ggplot(df, aes(
    x = sites,
    y = values,
    color = X,
    shape = Type,
    size = Type
  )) +
    geom_point(alpha = 1) +
    scale_shape_manual(values = c(1, 15), name=c("Values")) + 
    scale_size_manual(guide="none", values = size_point_graph) +
    guides(color = guide_legend(override.aes = list(size=point_size)),
           shape = guide_legend(override.aes = list(size=point_size))) +
    scale_color_manual(values = c(low,high),
                       name=c("Cell type"),
                       breaks=c("0", "1"),
                       labels=c("ES", "primary"))+

    labs(title = Title,
         x = "Sites",
         y = "Gaussian processes")+
    theme_minimal() +
    theme(
      text = element_text(size=textSize),
      plot.title = element_text(
        hjust = 0.5,
        color = colorTitle,
        size = sizeTitle,
        face = formeTitle
      ),
      axis.title.x = element_text(
        color = colorAxe,
        size = sizeAxe,
        face = formeAxe
      ),
      axis.title.y = element_text(
        color = colorAxe,
        size = sizeAxe,
        face = "bold"
      )
    )
  
p

```
